R code for 'Apparent physiological costs of early-life blood parasite infection in a vertebrate host

### 2022-02-12 MS_physiology_infection.Rmd
Add files via upload

---
title: "Manuscript_physiological_consequences_blood_parasite_infection"
author: "Tony Rinaud"
date: "`r Sys.Date()`"
output: pdf_document
---

# Analysis of blood parasite infection effect on physiological change of common buzzard nestlings:

  Data from common buzzard's (*Buteo buteo*) nestling of the years 2016, 2018, 2019 and 2020 which have been sampled twice during their nesting phase.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

@@

```{r packages, message=F, warning=F}
library(easypackages)
libraries('plotly','ggplot2','Hmisc','ggalt','plyr','chron','dplyr','stringr','tidyr','FSA','grid','gridExtra','ggsci','lme4','RVAideMemoire','lsmeans','MuMIn','EnvStats','devtools','RColorBrewer','tidyverse','brms','tidybayes','patchwork','pdftools','osmdata','rgdal','magrittr','purrr','forcats','modelr','ggstance','ggridges','cowplot','ggrepel','gganimate','ggridges','rptR','rdwd','sjPlot','sjlabelled','sjmisc','wesanderson', "ggpubr","ggsignif",'ggdist','ggeffect')

# Routine clean-up upon start
rm(list=ls())

theme_set(theme_tidybayes() + panel_border())

@@

```{r data, echo=F}

## Data source

data2016_2020=read.csv('./physiology_infection_stage_full_wideformat.csv', h=T, sep=';', dec=".")

### Setting classes of variables

data2016_2020$Nest=as.factor(data2016_2020$Nest)
data2016_2020$Date=chron(as.character(data2016_2020$Date), format="d/m/y")
data2016_2020$Infection_intensity=as.factor(data2016_2020$Infection_intensity)
data2016_2020$Prevalence=as.factor(data2016_2020$Prevalence)
data2016_2020$Breathing_rate=as.numeric(data2016_2020$Breathing_rate)
data2016_2020$Sex=as.factor(data2016_2020$Sex)
data2016_2020$Sampling_rank=as.factor(data2016_2020$Sampling_rank)
data2016_2020$Weight_g=scale(as.numeric(data2016_2020$Weight.2))
data2016_2020$Wing_cm=as.numeric(data2016_2020$Wing)
data2016_2020$body_temperature=as.numeric(data2016_2020$body_temperature)
data2016_2020$delta_body_condition=as.numeric(data2016_2020$delta_body_condition)
data2016_2020$Ambient_Temperature_AvgC=scale(as.numeric(data2016_2020$TemperatureAvgC))
data2016_2020$Treatment=as.factor(data2016_2020$Treatment)
data2016_2020$Year=as.factor(data2016_2020$Year)
data2016_2020$Growth_rate=as.numeric(data2016_2020$Growth_rate_unlog)
data2016_2020$days_resample=scale(data2016_2020$days_resample)
data2016_2020$Age_in_days=scale(as.numeric(data2016_2020$Age_in_days))

data2016_2020$Diff_Inf_Class=factor(data2016_2020$Diff_Inf_Class, levels=c("Uninfected","Increase","Peak","Decrease"))
data2016_2020=arrange(data2016_2020,data2016_2020$Diff_Inf_Class)

# Filtering non-applicable infection change

data2016_2020=data2016_2020[!is.na(data2016_2020$Diff_Inf_Class),]

>>>>

---
title: "R code for 'Apparent physiological costs of early-life blood parasite infection in a vertebrate host'"
author: "Tony Rinaud, Meinolf Ottensmann, Oliver Krüger & Nayden Chakarov"
output: html_document
---

This document provides all the `R code` used in our paper. Both the Rmarkdown file and the data can be downloaded from the accompanying GitHub repository on (<https://github.com/TonyRinaud/MS_physiology_infection>) as a zip archive containing all the files. We recommend to download or clone this [GitHub repository](https://github.com/TonyRinaud/MS_physiology_infection) in order to access the documentation together with all the files that are needed to repeat analyses shown in this document. Just click on the link above and then on the green box `Clone or download`. In order to function properly, the same structure of folders must be kept. If you have any questions, don't hesitate to contact [tony.rinaud\@gmail.com](mailto:tony.rinaud@gmail.com){.email}

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, strip.white = TRUE, collapse = TRUE, fig.align = "center",
                      warning = F, message = F, highlight = TRUE, tidy = TRUE)

# Routine clean-up upon start

rm(list=ls())
```

@@

A number of packages, all of which are available on CRAN are required to repeat analyses presented in this document. The code chunk below allows to load them for the analyses.

```{r packages, message=F, warning=F}
library(chron)
library(cowplot)
library(EnvStats)
library(lme4)
library(lsmeans)
library(ggalt)
library(ggdist)
library(ggeffects)
library(ggplot2)
library(ggpubr)
library(ggrepel)
library(ggridges)
library(ggsci)
library(ggsignif)
library(grid)
library(gridExtra)
library(magrittr)
library(patchwork)
library(rptR)
library(RVAideMemoire)
library(sjPlot)
library(sjlabelled)
library(sjmisc)
library(tidybayes)
library(tidyverse)
library(wesanderson)
theme_set(theme_tidybayes() + panel_border())
```

## Load raw data

The raw data is comprised of 276 common buzzard (*Buteo buteo*) individuals. Each individual was sampled twice over the course of nestling period during the years 2016 and 2018 - 2020. First, the raw data is loaded from the file `physiology_infection_stage_full_wideformat.csv` and formatted (scaling variables, reordering factor levels etc.) for downstream analyses

```{r data}
## Load data from csv file and code variable classes
## -----------------------------------------------------------------------------
df <- 
  read.csv('./physiology_infection_stage_full_wideformat.csv', h = T, sep = ';', dec = ".") %>% 
  mutate(Nest = as.factor(Nest)) %>% 
  mutate(Date = chron(as.character(Date), format = "d/m/y")) %>% 
  mutate(Infection_intensity.1 = as.factor(Infection_intensity.1)) %>% 
  mutate(Infection_intensity.2 = as.factor(Infection_intensity.2)) %>%
  mutate(Prevalence2 = as.factor(ifelse(Infection_intensity.2 == 0, 0, 1))) %>%
  mutate(Prevalence1 = as.factor(ifelse(Infection_intensity.1 == 0, 0, 1))) %>% 
  mutate(Breathing_rate = as.numeric(Breathing_rate)) %>% 
  mutate(Sex = as.factor(Sex)) %>% 
  mutate(Sampling_rank = as.factor(Sampling_rank)) %>% 
  mutate(Weight_g = scale(as.numeric(Weight.2))) %>% 
  mutate(Wing_cm = as.numeric(Wing.2)) %>% 
  mutate(body_temperature = as.numeric(body_temperature)) %>% 
  mutate(delta_body_condition = as.numeric(delta_body_condition)) %>% 
  mutate(Ambient_Temperature_AvgC = scale(as.numeric(TemperatureAvgC))) %>% 
  mutate(Treatment = as.factor(Treatment)) %>% 
  mutate(Year = as.factor(Year)) %>% 
  mutate(Growth_rate = as.numeric(Growth_rate)) %>% 
  mutate(days_resample = scale(days_resample)) %>% 
  mutate(Age_in_days = scale(as.numeric(Age_in_days))) %>% 
  mutate(Diff_Inf_Class = factor(Diff_Inf_Class, levels = c("Uninfected","Increase","Peak","Decrease")))

## Arrange data frame
## ----------------------------------------------------------------------------- 
df <- arrange(df, df$Diff_Inf_Class)

@@

## Effects on parasites

Investigating treatment effects on the prevalence of *Leucocytozoon* parasites:

```{r}
## Build contingency table: Infection prevalence and treatment: First sampling
## -----------------------------------------------------------------------------
pre_treatment <- with(df, table(Prevalence1, Treatment))
prop.test(pre_treatment)
pairwise.prop.test(pre_treatment)

## Build contingency table: Infection prevalence and treatment: Second sampling
## -----------------------------------------------------------------------------
post_treatment <- with(df, table(Prevalence2, Treatment))
prop.test(post_treatment)
pairwise.prop.test(post_treatment)

## Proportion of infected individuals across time points
## -----------------------------------------------------------------------------
infected_treatment <- 
  matrix(c(pre_treatment[2,], post_treatment[2,]), nrow = 2) %>% 
  set_colnames(., c("Pre", "Post")) %>% 
  set_rownames(., c("Control", "Malarone"))
prop.test(infected_treatment)
pairwise.prop.test(infected_treatment)

## Proportion of uninfected individuals across time points
## -----------------------------------------------------------------------------
uninfected_treatment <- 
  matrix(c(pre_treatment[1,], post_treatment[1,]), nrow = 2) %>% 
  set_colnames(., c("Pre", "Post")) %>% 
  set_rownames(., c("Control", "Malarone"))
prop.test(uninfected_treatment)
pairwise.prop.test(uninfected_treatment)

## Plotting count of infected / uninfected nestlings across treatments
## -----------------------------------------------------------------------------
tab_cont = data.frame(
  Count = c(infected_treatment[1,], uninfected_treatment[1,], infected_treatment[2,], uninfected_treatment[2,]),
  Timepoint = rep(c("Pre-treatment","Post-treatment"), 4),
  Treatment = rep(c("Control", "Treatment"), each = 4),
  Prevalence = rep(rep(c("Infected", "Uninfected"), each = 2),2)) %>% 
  mutate(Timepoint = factor(Timepoint, levels = c("Pre-treatment","Post-treatment"))) %>% 
  mutate(Prevalence = factor(Prevalence, levels = c("Uninfected","Infected")))

##  Create Figure 2a
## -----------------------------------------------------------------------------
treatment_prevalence_changes  <- ggplot(tab_cont) +  
  aes(Timepoint, Count, fill = Prevalence) +  
  facet_wrap(facets = "Treatment") +  
  geom_bar(stat = 'identity',  width = 0.3) +  
  labs(y = "Count of nestlings") +  
  geom_text(aes(label = Count),position  = position_stack(0.5)) +  
  scale_fill_manual(values = c("#D0D0D0","#A8A8A8")) +  
  theme(axis.title.x  =  element_blank(), 
        strip.text = element_text(size = 16, face = "bold"),
        legend.text = element_text(size = 14),
        legend.title = element_text(size = 16),
        axis.title.y = element_text(size = 18, margin =  unit(c(0, 3, 0, 0), "mm")),
        axis.text = element_text(size = 16))

## Plotting mean infection intensity values of nestlings across treatments
## -----------------------------------------------------------------------------
mean_infint = data.frame(
  Sampling_rank = factor(c(rep("1",length(df$Sampling_rank)),as.character(df$Sampling_rank)),levels = c("1","2"),
                         labels  =  c("Pre-treatment","Post-treatment")),
  Infection_intensity = as.numeric(c(as.character(df$Infection_intensity.1),as.character(df$Infection_intensity))),
  Treatment = factor(c(as.character(rep(df$Treatment,2))),levels  =  c("Control","Malarone")))

## Figure 2b
## -----------------------------------------------------------------------------
treatment_meaninf_changes = ggplot(mean_infint) +
  aes(Sampling_rank, Infection_intensity, group = Treatment) +
  geom_vline(xintercept  =  "Pre-treatment", linetype = "dashed", color = 'darkgrey') +
  geom_vline(xintercept = "Post-treatment", linetype = "dashed", color = 'darkgrey') +
  labs(y = "Mean infection intensity") +
  geom_point(aes(colour = Treatment),position = position_jitterdodge(jitter.height  =  0.1)) +
  stat_summary(fun = mean, geom = "line",size = 2,color = c("lightgrey","black","black","lightgrey")) +
  stat_summary(fun = mean, geom = "point",size = 5) +
  stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.04, size = 2) + 
  scale_color_manual(values = c("lightgrey","black")) + 
  theme_classic() + 
  theme(axis.title.x = element_blank(),
        legend.text = element_text(size = 16),
        legend.title = element_text(size = 16),
        axis.title.y = element_text(size = 18, margin =  unit(c(0, 3, 0, 0), "mm")),
        axis.text = element_text(size = 16),
        plot.margin = unit(c(0,0,0,30), "pt"))

## Panel wrapping Fig2 a & b:
## ----------------------------------------------------------------------------- 
Fig2 <- treatment_prevalence_changes + 
  treatment_meaninf_changes +
  plot_annotation(tag_levels = 'A') + 
  theme(plot.tag = element_text(size = 16, face = "bold"))
ggsave(Fig2,filename = "./Figs/efficiency_malarone.png",units = "px", width = 6000, height = 3000)

## Testing for difference in mean infection intensity between Pre and Post treatment
## -----------------------------------------------------------------------------
mean_before_after <- data.frame(
  Ring_no = 
    c(df$Ring_no[df$Sampling_rank %in% 2], df$Ring_no[df$Sampling_rank %in% 2]),
  Infection_intensity = c(as.character(df$Infection_intensity.1), as.character(df$Infection_intensity.2)),
  Treatment = 
    c(as.character(df$Treatment[df$Sampling_rank %in% 2]), as.character(df$Treatment[df$Sampling_rank %in% 2])),
  Timepoint = 
    c(rep("Pre-treatment", times = length(df$Treatment)), rep("Post-treatment", times = length(df$Infection_intensity.2)))) %>% 
  mutate(Treatment = as.factor(Treatment)) %>% 
  mutate(Timepoint = as.factor(Timepoint))

mal = mean_before_after[mean_before_after$Treatment %in% "Malarone",]
ctrl = mean_before_after[mean_before_after$Treatment %in% "Control",]
first = mean_before_after[mean_before_after$Timepoint %in% "Pre-treatment",]

## Pre-treatment difference in parasitic load between Treatments
## -----------------------------------------------------------------------------
perm.t.test(as.numeric(as.character(first$Infection_intensity)) ~ first$Treatment, progress = F)
sd(first$Infection_intensity[first$Treatment %in% "Control"])

## In malarone treated nestlings
## -----------------------------------------------------------------------------
perm.t.test(as.numeric(as.character(mal$Infection_intensity)) ~ mal$Timepoint, progress = F)

# In control nestlings
## -----------------------------------------------------------------------------
perm.t.test(as.numeric(as.character(ctrl$Infection_intensity))~ctrl$Timepoint, progress = F)
```

## Effects on physiology

Exploring effects on physiology of nestlings due to the treatment:

No treatment effect on breathing rate!

```{r}
lm_d1 = lm(Breathing_rate ~ Treatment, data = df)
summary(lm_d1)
```

No treatment effect on body temperature

```{r}
lm_d2 <- lm(body_temperature~Treatment, data = df)
summary(lm_d2)
```

No treatment effect on body condition change

```{r}
lm_d3 <- lm(delta_body_condition~Treatment, data = df)
summary(lm_d3)
```

No treatment effect on growth rates

```{r}
lm_d4 <- lm(Growth_rate~Treatment, data = df)
summary(lm_d4)

## Create Table S1
## -----------------------------------------------------------------------------
tab_model(
  lm_d1,lm_d2,lm_d3,lm_d4,
  show.reflvl = T,
  show.ci50 = T ,
  show.stat = T,
  string.stat = "t",
  pred.labels = c("Intercept","Treatment [Malarone]"),
  dv.labels = c("Breathing rate","Body temperature", "\U0394 Body condition","Growth rate"),
  file = "./Physiological_parameters_in_decreasing_control_treated.html")
```

## Infection stages as explanatory variable

Exploring differences in physiological measures that are associated with the infection stage (increase, peak or decreasing) of tested individuals.

```{r mixed model breathing rate, warning=F, message=F}
## Run model
## ----------------------------------------------------------------------------- 
lmer_br <- lmer(
  Breathing_rate ~ Diff_Inf_Class + Diff_Inf_Class:Treatment + 
    days_resample + Age_in_days + Weight_g + Sex + Ambient_Temperature_AvgC + Year + (1|Nest),
  data = df)

## Obtain predicted values for each infection stage
## -----------------------------------------------------------------------------
mod_br <- get_model_data(lmer_br, type = 'pred', terms = "Diff_Inf_Class") %>% 
   mutate(x = factor(c("Uninfected","Increase","Peak","Decrease"), 
                     levels = c("Uninfected","Increase","Peak","Decrease")))

## Extract sample size for infection stage
## -----------------------------------------------------------------------------
n <- table(df$Diff_Inf_Class[!is.na(df$Breathing_rate)])

## ad sample size as label
## -----------------------------------------------------------------------------
mod_br <- mod_br %>% 
  mutate(x = factor(c(paste0(x,"\nn= ",n)), levels = c(paste0(x,"\nn= ",n))))

## Plotting predicted values across infection stages
## -----------------------------------------------------------------------------
gg1_pred <- ggplot(mod_br) + 
  aes(x,predicted) + 
  geom_hline(yintercept = mean(mod_br$predicted, na.rm = T), linetype = "dashed") + 
  labs(x = "", y = "Breathing rate") + 
  ylim(c(min(mod_br$conf.low) - 1.5,max(mod_br$conf.high) + 1.5)) + 
  geom_errorbar(aes(ymin = mod_br$conf.low,ymax = mod_br$conf.high),colour = "black",size = 2, width = .1) + 
  geom_errorbar(aes(ymin = mod_br$predicted - mod_br$std.error, ymax = mod_br$predicted + mod_br$std.error),
                size = 8, colour = wes_palette("Zissou1",4), width = 0,) + 
  geom_point(color = "black", outlier.shape = NA, size = 2, width = .3) + 
  theme_classic() + 
  theme(legend.position = "none",
        axis.title = element_text(size = 14),
        axis.text = element_text(size = 12),
        axis.line.x = element_blank(), 
        axis.ticks.x = element_blank(), 
        plot.margin = margin(0, 0, 0, 0, "cm"), 
        plot.title = element_text(hjust = 0.5, face = "bold")) + 
  geom_segment(size = .8,
  aes(x = 1, y = max(mod_br$conf.high) + 1, xend = 4, yend = max(mod_br$conf.high) + 1)) + 
  geom_text(aes(x = 2.5, y = max(mod_br$conf.high) + 1.5),
  label = "n.s.", size = 5)
```

```{r mixed model body temp after, warning=FALSE, message=F}
## Model 
## -----------------------------------------------------------------------------
lmer_bt = lmer(
  body_temperature ~ Diff_Inf_Class + Diff_Inf_Class:Treatment + Ambient_Temperature_AvgC + 
    days_resample + Age_in_days + Weight_g + Sex + Year + (1|Nest),
  data = df)

## Predicted values for each infection stage
## -----------------------------------------------------------------------------
mod_bt = get_model_data(lmer_bt,type = 'pred',terms = "Diff_Inf_Class") %>%
  mutate(x = factor(c("Uninfected","Increase","Peak","Decrease"),
                    levels = c("Uninfected","Increase","Peak","Decrease")))

# Table with sample sizes specific for body temperature data
n2 = table(df$Diff_Inf_Class[!is.na(df$body_temperature)])

mod_bt <- mod_bt %>% 
  mutate(x = factor(c(paste0(x,"\nn= ",n2)), levels = c(paste0(x,"\nn= ",n2))))

# Plotting predicted values across infection stages
gg2_pred <- ggplot(mod_bt) + 
  aes(x,predicted) + 
  geom_hline(yintercept = mean(mod_bt$predicted, na.rm = T), linetype = "dashed") + 
  labs(x = "", y = "Body temperature") + 
  geom_errorbar(aes(ymin = mod_bt$conf.low,ymax = mod_bt$conf.high),colour = "black",size = 2, width = .1) + 
  geom_errorbar(aes(ymin = mod_bt$predicted - mod_bt$std.error,ymax = mod_bt$predicted + mod_bt$std.error),
                size = 8,colour = wes_palette("Zissou1",4), width = 0) + 
  geom_point(color = "black", outlier.shape = NA, size = 2, width = .3) + 
  theme_classic() + 
  theme(legend.position = "none",
        axis.title.y = element_text(size = 14, margin = unit(c(0,3,0,0),"mm")), 
        axis.text = element_text(size = 12),
        axis.line.x = element_blank(),
        axis.ticks.x = element_blank(), plot.margin = margin(0, 0, 0, 0, "cm"),
        plot.title = element_text(hjust = 0.5, face = "bold")) + 
  geom_segment(size = .8,
                 aes(x = 1, y = 40.6, xend = 4, yend = 40.6)) + 
    geom_text(aes(x = 2.5, y = 40.7),
              label = "n.s.", size = 5)
```

```{r mixed model body condition, warning=FALSE, message=F}
## Model 
## -----------------------------------------------------------------------------
lmer_bc <- lmer(
  delta_body_condition ~ Diff_Inf_Class+Diff_Inf_Class:Treatment+days_resample + 
    Age_in_days+Sex+Year+(1|Nest),
  data = df)

# Predicted values for each infection stage
mod_bc = get_model_data(lmer_bc,type = 'pred',terms = "Diff_Inf_Class") %>%
  mutate(x = factor(c("Uninfected","Increase","Peak","Decrease"), 
                  levels = c("Uninfected","Increase","Peak","Decrease")))

# Table with sample sizes specific for body condition change data
n3 = table(df$Diff_Inf_Class[!is.na(df$delta_body_condition)])

mod_bc <- mod_bc %>% 
  mutate(x = factor(c(paste0(x,"\nn= ",n3)), levels = c(paste0(x,"\nn= ",n3))))

gg3_pred = ggplot(mod_bc) + 
  aes(x,predicted) + 
  geom_hline(yintercept = mean(mod_bc$predicted, na.rm = T), linetype = "dashed") + 
  labs(x = "", y = "Δ Body condition") + 
  geom_errorbar(aes(ymin = mod_bc$conf.low,ymax = mod_bc$conf.high),colour = "black",size = 2, width = .1) + 
  geom_errorbar(aes(ymin = mod_bc$predicted - mod_bc$std.error,ymax = mod_bc$predicted + mod_bc$std.error),
                size = 8,colour = wes_palette("Zissou1",4), width = 0,) + 
  geom_point(color = "black", outlier.shape = NA, size = 2, width = .3) + 
  theme_classic() + 
  theme(legend.position = "none",
        axis.title = element_text(size = 14),
        axis.text = element_text(size = 12),
        axis.line.x = element_blank(),
        axis.ticks.x = element_blank(),
        plot.margin = margin(0, 0, 0, 0, "cm"),
        plot.title = element_text(hjust = 0.5, face = "bold")) + 
  geom_segment(size = .8, aes(x = 1, y = max(mod_bc$conf.high) + 17, xend = 4, yend = max(mod_bc$conf.high) + 17)) + 
  geom_text(aes(x = 2.5, y = max(mod_bc$conf.high) + 23),
              label = "n.s.", size = 5)
```

```{r mixed model body temp, warning=FALSE, message=F}
# Model 
lmer_gr = lmer(Growth_rate ~ Diff_Inf_Class + Diff_Inf_Class:Treatment + Sex + days_resample + Age_in_days + Year + (1|Nest), data= df)

# Predicted values for each infection stages
mod_gr = get_model_data(lmer_gr,type = 'pred',terms = "Diff_Inf_Class") %>%
  mutate(x = factor(c("Uninfected","Increase","Peak","Decrease"), levels = c("Uninfected","Increase","Peak","Decrease")))

# Table with sample sizes specific for growth rate data
n4 = table(df$Diff_Inf_Class[!is.na(df$Growth_rate)])

mod_gr <- mod_gr %>% 
  mutate(x = factor(c(paste0(x,"\nn= ",n4)), levels = c(paste0(x,"\nn= ",n4))))

gg4_pred = ggplot(mod_gr) + 
  aes(x,predicted) + 
  geom_hline(yintercept = mean(mod_gr$predicted, na.rm = T), linetype = "dashed") + 
  labs(x = "", y = "Growth rate") + 
  geom_errorbar(aes(ymin = mod_gr$conf.low,ymax = mod_gr$conf.high),colour = "black",size = 2, width = .1) + 
  geom_errorbar(aes(ymin = mod_gr$predicted - mod_gr$std.error,ymax = mod_gr$predicted + mod_gr$std.error), 
                size = 8,colour = wes_palette("Zissou1",4), width = 0,) + 
  geom_point(color = "black", outlier.shape = NA, size = 2, width = .3) + 
  theme_classic() + 
  theme(legend.position = "none", 
        axis.title.y = element_text(size = 14, margin = unit(c(0, 3, 0, 0), "mm")),
        axis.text = element_text(size = 12),
        axis.line.x = element_blank(),
        axis.ticks.x = element_blank(),
        plot.margin = margin(0, 0, 0, 30, "pt"),
        plot.title = element_text(hjust = 0.5, face = "bold")) + 
  geom_segment(size = .8, aes(x = 1, y = max(mod_gr$conf.high) + 1.2, xend = 4, yend = max(mod_gr$conf.high) + 1.2)) + 
    geom_text(aes(x = 2.5, y = max(mod_gr$conf.high) + 2),
              label = "n.s.", size = 5)
```

## Summarise models

The code below is used to summarise the model output and create Figure 4 and Table 3 as shown in the manuscript.

```{r}
## Figure 4 
## -----------------------------------------------------------------------------
fig4 <- gg1_pred + gg2_pred + gg3_pred + gg4_pred + plot_annotation(tag_levels = 'A') & 
  theme(plot.tag = element_text(size = 16, face = "bold"))
ggsave(fig4,filename = "./Figs/predicted_physio_param_infection_stages.png",units = "px", width = 3700, height = 3700)

# Table 3
## -----------------------------------------------------------------------------
sjPlot::tab_model(
  lmer_br,lmer_bt,lmer_bc,lmer_gr,
  show.ci50 = T ,
  show.stat = T,
  string.stat = "t",
  df.method = "kr",
  dv.labels = c("Breathing rate","Body temperature","\U0394 Body condition","Growth rate"),
  file = "Table_results_lmer_physiological_parameters_p.val_kr.html")
```

## Parasitemia as explanatory variable

Exploration of effects by parasitemia levels on physiology of individuals.

```{r mixed model breathing rate 2, warning=F, message=F}
## Model 
## -----------------------------------------------------------------------------
lmer_br_para = lmer(
  Breathing_rate ~ Infection_intensity.2 + Infection_intensity.2:Treatment + days_resample + 
    Age_in_days + Weight_g + Sex + Ambient_Temperature_AvgC + Year + (1|Nest), data= df)

mod_br_para=get_model_data(lmer_br_para,type = 'pred',terms = "Infection_intensity.2") %>% 
   mutate(x = factor(c("0","1","2","3","4"), levels = c("0","1","2","3","4")))

m1 = table(df$Infection_intensity.2[!is.na(df$Breathing_rate)])

mod_br_para <- mod_br_para %>% 
  mutate(x = factor(c(paste0(x,"\nn= ",m1)), levels = c(paste0(x,"\nn= ",m1))))
mod_br_para$lower_CI= c(31.11, mod_br_para$predicted[1] - 2.96, mod_br_para$predicted[1] - 6.49,mod_br_para$predicted[1] - 4.67, mod_br_para$predicted[1] - 4.63)
mod_br_para$higher_CI= c(41.42,mod_br_para$predicted[1]+3.90,mod_br_para$predicted[1]+1.75,mod_br_para$predicted[1]+1.46,mod_br_para$predicted[1]+4.45)
mod_br_para$se= c(2.61,1.74,2.09,1.56,2.31)

gg1_pred_para = ggplot(mod_br_para) + 
  aes(as.factor(mod_br_para$x),mod_br_para$predicted) + 
  geom_hline(yintercept = mean(mod_br_para$predicted, na.rm = T), linetype = "dashed") + 
  labs(x = "", y = "Breathing rate") + 
  geom_errorbar(aes(ymin = mod_br_para$lower_CI,ymax = mod_br_para$higher_CI),colour = "black",size = 2, width = .1) + 
  geom_errorbar(aes(ymin = mod_br_para$predicted - mod_br_para$se, ymax = mod_br_para$predicted + mod_br_para$se),
                size = 8, colour = wes_palette("Zissou1",5), width = 0,) + 
  geom_point(color = "black", outlier.shape = NA, size = 4, width = .3) + 
  theme_classic() + 
  theme(legend.position = "none",
        axis.title = element_text(size = 14),
        axis.text = element_text(size = 12),
        axis.line.x = element_blank(),
        axis.ticks.x = element_blank(), plot.margin = margin(0, 0, 0, 0, "cm"),
        plot.title = element_text(hjust = 0.5, face = "bold")) + 
  geom_segment(size = .8, aes(x = 1, y = max(mod_br_para$higher_CI, na.rm = T) + 1.2,
                              xend=5, yend = max(mod_br_para$higher_CI, na.rm = T) + 1.2)) + 
    geom_text(aes(x = 3, y = max(mod_br_para$higher_CI, na.rm = T) + 2),
              label = "n.s.", size = 5)
```

```{r  mixed model body temperature 2, warning=FALSE, message=F, echo=T}
# Model

lmer_bt_para = lmer(
  body_temperature ~ Infection_intensity.2 + Infection_intensity.2:Treatment + days_resample +
    Ambient_Temperature_AvgC + Age_in_days + Weight_g + Sex + Year + (1|Nest), data= df)

mod_bt_para = get_model_data(lmer_bt_para, 
                             type = 'pred',
                             terms = "Infection_intensity.2") %>% 
   mutate(x = factor(c("0","1","2","3","4"), levels = c("0","1","2","3","4")))

m2 = table(df$Infection_intensity.2[!is.na(df$body_temperature)])

mod_bt_para <- mod_bt_para %>% 
  mutate(x = factor(c(paste0(x,"\nn= ",m2)), levels = c(paste0(x,"\nn= ",m2))))
mod_bt_para$lower_CI= c(39.40,mod_bt_para$predicted[1] - 0.40,mod_bt_para$predicted[1] - 0.60,mod_bt_para$predicted[1] - 0.65,mod_bt_para$predicted[1] - 1.38)
mod_bt_para$higher_CI= c(40.40,mod_bt_para$predicted[1]+0.55,mod_bt_para$predicted[1]+0.36,mod_bt_para$predicted[1]+0.19,mod_bt_para$predicted[1]+(-0.24))
mod_bt_para$se= c(0.23,0.24,0.24,0.21,0.29)

gg2_pred_para = ggplot(mod_bt_para) + 
  aes(as.factor(mod_bt_para$x),mod_bt_para$predicted) + 
  geom_hline(yintercept = mean(mod_bt_para$predicted, na.rm = T), linetype = "dashed") + 
  geom_errorbar(aes(ymin = mod_bt_para$lower_CI,ymax = mod_bt_para$higher_CI),colour = "black",size = 2, width = .1) + 
  geom_errorbar(aes(ymin = mod_bt_para$predicted - mod_bt_para$se,ymax = mod_bt_para$predicted + mod_bt_para$se),
                size =8,colour = wes_palette("Zissou1",5), width = 0,) + 
  labs(x = "", y = "Body temperature") + 
  geom_point(color = "black", outlier.shape = NA, size =4, width = .3) + 
  theme_classic() + 
  theme(legend.position = "none",
        axis.title = element_text(size = 14),
        axis.text = element_text(size = 12),
        axis.line.x = element_blank(), 
        axis.ticks.x = element_blank(), plot.margin = margin(0, 0, 0, 0, "cm"),
        plot.title = element_text(hjust = 0.5, face = "bold")) + 
  geom_segment(size = .8, aes(x = 1, y = max(mod_bt_para$higher_CI, na.rm = T) + 0.2,
                              xend = 5, yend = max(mod_bt_para$higher_CI, na.rm = T) + 0.2)) + 
    geom_text(aes(x = 3, y = max(mod_bt_para$higher_CI, na.rm = T) + 0.25),
              label = "**", size = 8) + 
  geom_segment(size = .8,
                 aes(x = 1, y = max(mod_bt_para$higher_CI, na.rm = T) + 0.2, xend=1, yend = max(mod_bt_para$higher_CI, na.rm = T) + 0.05)) + 
  geom_segment(size = .8,
                 aes(x =5, y = max(mod_bt_para$higher_CI, na.rm = T) + 0.05, xend=5, yend = max(mod_bt_para$higher_CI, na.rm = T) + 0.2))
```

```{r mixed model body condition change 2, warning=FALSE, message=F, echo=T}
# Model 
lmer_bc_para <- lmer(
  delta_body_condition ~ Infection_intensity.2 + Infection_intensity.2:Treatment + days_resample +
    Age_in_days + Sex + Year + (1|Nest), data= df)

mod_bc_para=get_model_data(lmer_bc_para,type = 'pred',terms = "Infection_intensity.2") %>% 
   mutate(x = factor(c("0","1","2","3","4"), levels = c("0","1","2","3","4")))

m3 = table(df$Infection_intensity.2[!is.na(df$delta_body_condition)])

mod_bc_para <- mod_bc_para %>% 
  mutate(x = factor(c(paste0(x,"\nn= ",m3)), levels = c(paste0(x,"\nn= ",m3))))
mod_bc_para$lower_CI= c(-65.99,mod_bc_para$predicted[1] - 5.71,mod_bc_para$predicted[1] - 53.27,mod_bc_para$predicted[1] - 37.05,mod_bc_para$predicted[1] - 71.88)
mod_bc_para$higher_CI= c(25.59,mod_bc_para$predicted[1]+51.67,mod_bc_para$predicted[1]+16.43,mod_bc_para$predicted[1]+16.17,mod_bc_para$predicted[1]+4.09)
mod_bc_para$se= c(23.20,14.57,17.69,13.51,19.29)

gg3_pred_para = ggplot(mod_bc_para) + 
  aes(as.factor(mod_bc_para$x),mod_bc_para$predicted) + 
  geom_hline(yintercept = mean(mod_bc_para$predicted, na.rm = T), linetype = "dashed") + 
  labs(x = "", y = "Δ Body condition") + 
  geom_errorbar(aes(ymin = mod_bc_para$lower_CI,ymax = mod_bc_para$higher_CI),colour = "black",size = 2, width =.1) + 
  geom_errorbar(aes(ymin = mod_bc_para$predicted-mod_bc_para$se,ymax = mod_bc_para$predicted+mod_bc_para$se),size =8,colour = wes_palette("Zissou1",5), width = 0,) + 
  geom_point(color = "black", outlier.shape = NA, size =4, width = .3) + 
  theme_classic() + 
  theme(legend.position = "none",
        axis.title = element_text(size = 14),  
        axis.text = element_text(size = 12), 
        axis.line.x = element_blank(),
        axis.ticks.x = element_blank(), plot.margin = margin(0, 0, 0, 0, "cm"),
        plot.title = element_text(hjust = 0.5, face = "bold")) + 
  geom_segment(size = .8,
                 aes(x = 1, y =-105, xend=5, yend =-105)) + 
    geom_text(aes(x =3, y =-110),
              label= "p=0.08", size = 5) + 
  geom_segment(size = .8,
                 aes(x = 1, y =-105, xend=1, yend =-105+5)) + 
  geom_segment(size = .8,
                 aes(x =5, y =-105, xend=5, yend =-105+5))
```

```{r mixed model growth rate 2, warning=FALSE, message=F, echo=T}
# Model 
lmer_gr_para = lmer(
  Growth_rate ~ Infection_intensity.2 + Infection_intensity.2:Treatment +
    days_resample + Sex + Age_in_days + Year + (1|Nest), data= df)

# Predicted values for post-treament parasitemia
mod_gr_para=get_model_data(lmer_gr_para,type = 'pred',terms = "Infection_intensity.2") %>% 
   mutate(x = factor(c("0","1","2","3","4"), levels = c("0","1","2","3","4")))

m4 = table(df$Infection_intensity.2[!is.na(df$Growth_rate)])

mod_gr_para <- mod_gr_para %>% 
  mutate(x = factor(c(paste0(x,"\nn= ",m4)), levels = c(paste0(x,"\nn= ",m4))))
mod_gr_para$lower_CI= c(11.44,mod_gr_para$predicted[1] - 2.22,mod_gr_para$predicted[1] - 10.47,mod_gr_para$predicted[1] - 6.60,mod_gr_para$predicted[1] - 11.39)
mod_gr_para$higher_CI= c(25.14,mod_gr_para$predicted[1]+7.08,mod_gr_para$predicted[1]+0.92,mod_gr_para$predicted[1]+1.91,mod_gr_para$predicted[1]+0.89)
mod_gr_para$se= c(3.47,2.36,2.89,2.16,3.12)
